---
title: "Assignment 2b"
author: "Yinqi Zhao"
date: "2/23/2020"
output: pdf_document
---

# Secant Method

## 1. Conducting the Secant method
Here is a function which will conduct the Secant method.
```{r}
# secant method
secant <- function(fx, start0, start1, tol = 1e-3, maxitr = 1000){
  res <- c(start0, start1)
  distance <- -Inf
  x0 <- start0
  x1 <- start1
  i <- 1
  while(i <= maxitr & abs(distance) > tol){
    x2 <- x1 - fx(x1) * (x1 - x0) / (fx(x1) - fx(x0))
    res <- c(res, x2)
    distance <- abs(x2 - x1)
    x0 <- x1
    x1 <- x2
    print(paste0("iteration ", i, ": ", "X = ", x2, ", Y = ", fx(x2)))
    i <- i + 1
  }
  if(abs(distance) > tol){
    print("fail to converge, please try another start value")
  }
  print(paste0("converged! the root is ", x1))
  return(list(root = x2,
              res = res))
}
```

We test it on 2 functions, the first one is $f(x) = \cos x - x$, with start value $x_0 = 1$ and $x_1 = 2$.
```{r}
f1 <- function(x){
  cos(x) - x
}

res.secant1 <- secant(fx = f1, start0 = 1, start1 = 2)
```

The second function is $f(x) = \log x - \exp(x)$, with the same start value as above.
```{r}
f2 <- function(x){
  log(x) - exp(-x)
}

res.secant2 <- secant(fx = f2, start0 = 1, start1 = 2)
```

## 2. Comparing to the Newtons method.
A function to conduct the Newton's method.
```{r}
newton <- function(fx, fx2, start, tol = 1e-3, maxitr = 1000){
  res <- start
  distance <- -Inf
  x1 <- start
  i <- 1
  while(i <= maxitr & abs(distance) > tol){
    x2 <- x1 - fx(x1) / fx2(x1)
    res <- c(res, x2)
    distance <- abs(x2 - x1)
    x1 <- x2
    print(paste0("iteration ", i, ": ", "X = ", x1, ", Y = ", fx(x1)))
    i <- i + 1
  }
  if(abs(distance) > tol){
    print("fail to converge, please try another start value")
  }
  print(paste0("converged! the root is ", x1))
  return(list(root = x2,
              res = res))
}

f1.d <- function(x){
  -sin(x) - 1
}

f2.d <- function(x){
  1 / x + exp(-x)
}
```

Choose 0 as the start point of Newton's method.
```{r}
invisible(capture.output(res.newton1 <- newton(f1, f1.d, start = 1)))
invisible(capture.output(res.newton2 <- newton(f2, f2.d, start = 1)))

results <- data.frame(Secant.itr = c(length(res.secant1$res), length(res.secant2$res)),
                      Secant.root = c(res.secant1$root, res.secant2$root),
                      Newton.itr = c(length(res.newton1$res), length(res.newton2$res)),
                      Newton.root = c(res.newton1$root, res.newton2$root))
row.names(results) <- c("funtion1", "function2")
results
```
Both root finding method produce the same results. It is not very straightforward to compare the convergence speed of the two method. Based on the initial value of 1 for Newton's method and 1, 2 for Secant method, Newton's method is a little bit faster than the Secant but the difference is not very big. Both method converge very fast


# Plot the process of Secant method
```{r}
plot.process <- function(fx, res, xrange = c(-1, 1)){
  grid <- seq(from = xrange[1], to = xrange[2], length.out = 1000)
  plot(grid, fx(grid), type = "l", xlab = "x", ylab = "y")
  abline(a = 0, b = 0)
  for(i in 1:length(res)){
    x <- res[i]
    segments(x0 = x, y0 = 0, x1 = x, y1 = fx(x), col = "red")
    if(i > 1){
      segments(x0 = x, y0 = fx(x), x1 = res[i - 1], y1 = fx(res[i - 1]), col = "red", lty = 5)
    }
  }
}

plot.process(f1, res.secant1$res, xrange = c(0, 3))
plot.process(f2, res.secant2$res, xrange = c(0.5, 2.5))
```

